---
title: "US Brewery and Beer Case Study"
author: "Dan Crouthamel, Pradeep Pradeepkumar"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(stats)
library('knitr')
pagebreak <- function() {
  if(knitr::is_latex_output())
    return("\\newpage")
  else
    return('<div style="page-break-before: always;" />')
}
```

## Case Study Introduction
Our client, Beer City, in Pennsylvania, is looking to expand their offering of domestic beer products. Using a recent list of 558 breweries and 2410 different beer products, we will assess some current interesting attributes about this landscape. This includes.

*  Number of Breweries per State
*  Median Alcohol Content per State (ABV)
*  International Bitterness Unit per State (IBU)
*  States with the highest ABV and IBU
*  Statistical summary on ABV and the relationship between ABV and IBU
*  Fun examination of highest ABV with low IBU

### Breweries per State

```{r brewPlot, fig.align='center', fig.width=11}
brewData <- read.csv("Breweries.csv", header = TRUE)

# brewSummary <- summary(brewData$State)
# brewSummary

brewTable <- brewData %>% count(State)
names(brewTable) <- c("State","Count")
# brewTable

brewPlot <-ggplot(data=brewTable, aes(x=reorder(State, -Count), y=Count, fill=State))
brewPlot +
  geom_bar(stat="identity") +
  ylab("Brewery Count") +
  xlab("State") +
  ggtitle("Brewery Count in the United States") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
```

The above graphic outlines the number of breweries in each state, with Colorado and California having the most.

### Read Beer data, then merge with Brewery data

```{r}
# Read beer data, then merge with Brewery data
beerData <- read.csv("Beers.csv")

# Clean up some columns so that things look nice when joining the tables
brewData %<>% rename("Brewery Name" = "Name")
beerData %<>% rename("Beer Name" = "Name")
beerData %<>% rename("Brew_ID" = "Brewery_id")

# Using inner_join of dplyr, could have used merged too, but I like this better
mergedData <- inner_join(brewData, beerData, by = c("Brew_ID" = "Brew_ID"))
kable(head(mergedData))
kable(tail(mergedData))
```

A sample of the beginning and ending of the data set appears above and includes the following observations.

* Brew_ID
* Brewery Name
* City
* State
* Beer Name
* Beer_ID
* ABV
* IBU
* Style
* Ounces

### Report the number of NA's in each column

```{r}
kable(colSums(is.na(mergedData)))
```

The above graphic shows us that we have 62 NAs in the ABV column, and 1005 NAs in the IBU column.

### Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

```{r fig.align='center', fig.width=11}

# Medians
medianABV <- mergedData %>% group_by(State) %>% summarise(median = median(ABV, na.rm = TRUE))
medianIBU <- mergedData %>% group_by(State) %>% summarise(median = median(IBU, na.rm = TRUE))

medianABV <- mergedData %>% group_by(State) %>% summarise(median = mean(ABV, na.rm = TRUE))
medianIBU <- mergedData %>% group_by(State) %>% summarise(median = mean(IBU, na.rm = TRUE))

abvPlot <-ggplot(data=medianABV, aes(x=reorder(State,-median), y=median, fill=State))
abvPlot +
  geom_bar(stat="identity") +
  ylab("Median ABV") +
  xlab("State") +
  ggtitle("Median ABV (Alcohol By Volume) per State") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "top")

ibuPlot <-ggplot(data=medianIBU, aes(x=reorder(State, -median), y=median, fill=State))
ibuPlot +
  geom_bar(stat="identity") +
  ylab("Median IBU") +
  xlab("State") +
  ggtitle("Median IBU (International Bitterness Units) per State") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "top")
```

There is no IBU data for South Dakota.

### Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?


```{r}
# Max Stuff
maxABV <- mergedData %>% group_by(State) %>% summarise(max = max(ABV, na.rm = TRUE))
maxABV.row <- maxABV[which.max(maxABV$max),]
maxABV.row

maxIBU <- mergedData %>% group_by(State) %>% summarise(max = max(IBU, na.rm = TRUE))
maxIBU.row <- maxIBU[which.max(maxIBU$max),]
maxIBU.row
```

Colorado has the highest ABV and Oregon has the highest IBU

### Summary statistics for the ABV variable.

```{r}
summary(mergedData$ABV)
#summary(mergedData$ABV[!is.na(mergedData$ABV)])
```

There are 62 cases with no ABV data, which could be non-alcoholic beer. Maybe. The lowest is .001, which is considered non alcoholic and the highest is 12.8%, which is quite high for a beer. 50% of the data lies between 5 and 6.7 % ABV, which seems typical of beer.

### Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.

```{r fig.align='center', fig.width=11}
mergedData1 <- na.omit(mergedData) 
ggplot(mergedData1[!(is.na(mergedData1$IBU)|is.na(mergedData1$ABV)),], aes(x=IBU, y=ABV)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = TRUE)
summary(lm(mergedData1$ABV ~ mergedData1$IBU))
```

Above we can see somewhat a linear trend between ABV and IBU. Something the should be considered though is the concept of relative bitterness vs perceived bitterness. For example, a beer that has 60 IBUs and is 6% ABV will taste about as bitter as a 90 IBU beer with 9% ABV. A higher ABV may indicate more malts, which is sweeter and will balance the precevied bitterness. So although there appears to be a weak linear trend, the average perceived bitterness could be flat (constant).

### Find a beer with maximum ABV and IBU less than 50

``` {r echo=TRUE}
ibu_nona <- mergedData[!is.na(mergedData$IBU),]
ibu_less_50 <- ibu_nona[ibu_nona$IBU < 50,]
max_abv <- head(ibu_less_50[order(ibu_less_50$ABV, decreasing = TRUE),],1)
name <- max_abv$`Beer Name`
```

Beer named  `r name`  has highest alcohol content at `r max_abv$ABV` and IBV value at `r max_abv$IBU`
